Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-14 manuscript.Rmd
even more NA fixes

timelines <- function(ds) {
  ds$id_uni <- as.numeric(factor(ds$id_uni))
  lims <- as_hms(c('09:00:00', '21:59:00'))
  hour_breaks = as_hms(c('09:00:00', '10:00:00', '11:00:00', '12:00:00', '13:00:00', '14:00:00', '15:00:00', '16:00:00', '17:00:00', '18:00:00', '19:00:00', '20:00:00', '21:00:00'))
  label_breaks = c("9am","","11am","","1pm","","3pm","","5pm","","7pm","","9pm")
  ggplot(ds, aes(x = clock_time_start, y = prop, fill = Position)) + 
  geom_bar(stat = "identity") + 
  facet_wrap("id_uni",ncol = 1, strip.position = "left") + 
  scale_fill_manual(values = c("Nap" = "gray",pal), name = "") + ylab("") + 
  scale_x_time(breaks = hour_breaks, labels = label_breaks, name = "", limits = lims) + 
  theme(legend.position = "bottom",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 11),
        strip.text.y.left = element_text(angle = 360))
  #https://ggplot2tor.com/scales/scale_x_time for options
  # strip.text.y = element_blank()
  # strip.background = element_blank(),
}

>>>>

timelines <- function(ds) {
  ds$id_uni <- as.numeric(factor(ds$id_uni))
  ds <- ds %>% mutate(prop = na_if(prop, 0))
  lims <- as_hms(c('09:00:00', '21:59:00'))
  hour_breaks = as_hms(c('09:00:00', '10:00:00', '11:00:00', '12:00:00', '13:00:00', '14:00:00', '15:00:00', '16:00:00', '17:00:00', '18:00:00', '19:00:00', '20:00:00', '21:00:00'))
  label_breaks = c("9am","","11am","","1pm","","3pm","","5pm","","7pm","","9pm")
  ggplot(ds, aes(x = clock_time_start, y = prop, fill = Position)) + 
  geom_bar(stat = "identity") + 
  facet_wrap("id_uni",ncol = 1, strip.position = "left") + 
  scale_fill_manual(values = c("Nap" = "gray",pal), name = "") + ylab("") + 
  scale_x_time(breaks = hour_breaks, labels = label_breaks, name = "", limits = lims) + 
  theme(legend.position = "bottom",
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 11),
        strip.text.y.left = element_text(angle = 360))
  #https://ggplot2tor.com/scales/scale_x_time for options
  # strip.text.y = element_blank()
  # strip.background = element_blank(),
}

### 2023-03-14 manuscript.tex
even more NA fixes

\begin{tabular}{lllll}
\toprule
 & \multicolumn{2}{c}{Group} & \multicolumn{2}{c}{Individual} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Position & Younger & Older & Younger & Older\\
\midrule
Upright & 7.7\% (9.3) & 18.6\% (7.4) & 10.4\% (13.6) & 18.7\% (8.4)\\
Sitting & 24.9\% (11.5) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
Prone & 14.4\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
Held & 12.7\% (7.1) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\
\bottomrule
\end{tabular}

@@

The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In Franchak (2019), upright time increased from 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months. Similarly, younger infants' sitting time estimated from full-day group models was 7.68\% compared with 18.57\% for the older group. In Franchak (2019), prone time increased ranged from 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months. In the current study, younger infants' prone time was was 14.38\% and 14.45\% for older infants. Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting depending on the specifics. Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants. Sitting was greater for older infants (44.42\%) compared with younger infants (24.92\%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

Figure \ref{fig:age} shows that both group and individual model predictions captured age-related differences in body position. Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.50, \emph{p} = .005). Similarly, age predicted increases in sitting time using group (\emph{r} = 0.70, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001). In contrast, age predicted decreases in supine time using group (\emph{r} = -0.66, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001). No clear age trends were found for prone time. Age was not significantly related to held time for group models (\emph{r} = -0.22, \emph{p} = .212), but individual models showed a significant negative correlation (\emph{r} = -0.42, \emph{p} = .021) resulting from an outlier in the younger group.

>>>>

\begin{tabular}{lllll}
\toprule
 & \multicolumn{2}{c}{Group} & \multicolumn{2}{c}{Individual} \\
\cmidrule(r){2-3} \cmidrule(r){4-5}
Position & Younger & Older & Younger & Older\\
\midrule
Upright & 7.7\% (9.3) & 18.6\% (7.4) & 10.4\% (13.6) & 18.7\% (8.4)\\
Sitting & 25.4\% (12.4) & 44.4\% (10.1) & 18.8\% (16.3) & 46.9\% (13.3)\\
Prone & 14.5\% (13.8) & 14.4\% (6.0) & 12.5\% (10.2) & 16.9\% (10.6)\\
Supine & 38.5\% (24.0) & 14.0\% (8.4) & 40.2\% (31.6) & 10.0\% (9.5)\\
Held & 13.1\% (7.4) & 8.5\% (5.4) & 17.1\% (20.8) & 7.4\% (7.6)\\
\bottomrule
\end{tabular}

@@

The most straightforward comparisons between the two investigations are upright and prone, because they were defined identically. In Franchak (2019), upright time increased from 0.6-5.5\% of the time for infants 3-6 months, increasing to 22.0\% at 12 months. Similarly, younger infants' sitting time estimated from full-day group models was 7.72\% compared with 18.57\% for the older group. In Franchak (2019), prone time increased ranged from 2.9-9.2\% of the time for infants 3-6 months, and was 7.2\% at 12 months. In the current study, younger infants' prone time was was 14.46\% and 14.45\% for older infants. Held, sitting, and supine are more difficult to compare because prior work (Franchak, 2019) considered held any time that infants were in caregivers' arms or on their laps, however, we only counted laying or sitting on caregivers' laps as supine and sitting, respectively. Further complicating the comparison, Franchak (2019) included a reclined category that we would classify as either supine or sitting depending on the specifics. Although the numbers are less directly comparable, our results are similar in showing that held and supine time was less for older infants than younger infants. Sitting was greater for older infants (44.42\%) compared with younger infants (25.37\%). Moreover, sitting was the most frequent body position followed by upright in 12-month-olds (Franchak, 2019), just as we observed in the current study. Overall, the estimates from the current investigation are consistent with age differences found using a different methodology.

Figure \ref{fig:age} shows that both group and individual model predictions captured age-related differences in body position. Age in months and upright time were significantly, positively correlated using group (\emph{r} = 0.59, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.50, \emph{p} = .005). Similarly, age predicted increases in sitting time using group (\emph{r} = 0.68, \emph{p} \textless{} .001) and individual models (\emph{r} = 0.72, \emph{p} \textless{} .001). In contrast, age predicted decreases in supine time using group (\emph{r} = -0.66, \emph{p} \textless{} .001) and individual models (\emph{r} = -0.61, \emph{p} \textless{} .001). No clear age trends were found for prone time. Age was not significantly related to held time for group models (\emph{r} = -0.23, \emph{p} = .200), but individual models showed a significant negative correlation (\emph{r} = -0.42, \emph{p} = .021) resulting from an outlier in the younger group.
